---
title: "Santander Cycle Bikesharing in London"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(readxl)
library(janitor)
library(jsonlite)
library(httr)
library(lubridate)
library(rvg)
library(dplyr)
library(sf)
library(tmap)
library(plotly)

# Importing Bike Rental Station locations from API

r <- GET("https://api.tfl.gov.uk/BikePoint/")

data = fromJSON(rawToChar(r$content))
names(data)
data$id

bikeStations <- data %>% 
  separate(id, into=c("bptext","StationID"), convert = TRUE, sep = "_")

# Importing the whole bikerental dataset

#load("C:/Users/konra/tubCloud/Uni/20WiSe/DataScience/MidsemesterAssignment/final project/Rohdaten/BikeExport.RData")
#load("C:/Users/konra/tubCloud/Uni/20WiSe/DataScience/MidsemesterAssignment/final project/Rohdaten/allroutes.RData")

load("C:/Users/loren/tubCloud/Shared/MidsemesterAssignment/final project/Rohdaten/BikeExport.RData")
load("C:/Users/loren/tubCloud/Shared/MidsemesterAssignment/final project/Rohdaten/allroutes.RData")

# Getting only bike 15776

bikestops_clean <- data %>%
  separate(col = id, sep = "_", into = c(NA, "id")) %>% 
  select(id, commonName, lat, lon) %>% 
  arrange(id) %>% 
  mutate(id = as.numeric(id))

bike_15776 <- bikerawdata %>% 
  filter(Bike.Id == 15776) %>%
  mutate(EndDate = strptime(End.Date, "%d/%m/%Y %H:%M")) %>% 
  arrange(EndDate) %>% 
  inner_join(bikestops_clean, by = c("StartStation.Id" = "id")) %>% 
  rename(commonName_start = commonName, lat_start = lat, lon_start = lon) %>% 
  inner_join(bikestops_clean, by = c("EndStation.Id" = "id")) %>% 
  rename(commonName_end = commonName, lat_end = lat, lon_end = lon)

allroutes <- allroutes %>% 
  left_join(bike_15776, by = c("id" = "Rental.Id")) %>% 
  mutate(Starttime = lubridate::hour(strptime(Start.Date, "%d/%m/%Y %H:%M")) +
           lubridate::minute(strptime(Start.Date, "%d/%m/%Y %H:%M")) / 60,
         dayride = floor(Starttime) > 7 & floor(Starttime) < 20)

# Using a little bit of it to make it faster at first

bikerawdata <- bikerawdata %>%
  mutate(EndDate = strptime(`End.Date`, "%d/%m/%Y %H:%M"), EndDate2 = format(EndDate, "%d/%m/%Y")) %>%
  mutate(StartDate = strptime(`Start.Date`, "%d/%m/%Y %H:%M"), StartYear = format(StartDate, "%Y"), StartMonth = format(StartDate, "%m"))

bikes062019 <- bikerawdata %>%
  filter(StartMonth == "06" & StartYear == "2019") %>%
  mutate(convTravelTime = (as.numeric(EndDate) - as.numeric(StartDate)) / 60) %>%
  mutate(StationID = StartStation.Id) %>%
  inner_join(bikeStations) %>%
  mutate(startLat = lat) %>%
  mutate(startLon = lon) 

bikes062020 <- bikerawdata %>%
  filter(StartMonth == "06" & StartYear == "2020") %>%
  mutate(convTravelTime = (as.numeric(EndDate) - as.numeric(StartDate)) / 60) %>%
  mutate(StationID = StartStation.Id) %>%
  inner_join(bikeStations) %>%
  mutate(startLat = lat) %>%
  mutate(startLon = lon)

# London map

#LondonWard <- st_read("C:/Users/konra/tubCloud/Uni/20WiSe/DataScience/MidsemesterAssignment/Rohdaten/Boroughs_London/London_Borough_Excluding_MHW.shp") %>%
#  st_transform(4326)
#LondonLSOA <- st_read("C:/Users/konra/tubCloud/Uni/20WiSe/DataScience/MidsemesterAssignment/Rohdaten/Boroughs_London/LSOA_2011_London_gen_MHW.shp") %>% 
#  st_transform(4326)

LondonWard <- st_read("C:/Users/loren/tubCloud/Shared/MidsemesterAssignment/Rohdaten/Boroughs_London/London_Borough_Excluding_MHW.shp") %>%
  st_transform(4326)
LondonLSOA <- st_read("C:/Users/loren/tubCloud/Shared/MidsemesterAssignment/Rohdaten/Boroughs_London/LSOA_2011_London_gen_MHW.shp") %>% 
  st_transform(4326)

startpoints <- st_as_sf(bikes062019, coords = c("lon", "lat"), crs = 4326)

sum_startpoints <- startpoints %>%
  group_by(StartStation.Id, StartStation.Name) %>%
  summarise(stationcount = n()) %>%
  relocate(StartStation.Name, .before = StartStation.Id)

startpoints2 <- st_as_sf(bikes062020, coords = c("lon", "lat"), crs = 4326)

sum_startpoints2 <- startpoints2 %>%
  group_by(StartStation.Id, StartStation.Name) %>%
  summarise(stationcount = n()) %>%
  relocate(StartStation.Name, .before = StartStation.Id)

starts_joinedLSOA <- startpoints %>%
  st_join(LondonLSOA, join = st_within)

starts_joinedLSOA2 <- startpoints2 %>%
  st_join(LondonLSOA, join = st_within)

# Counting residents

LSOAcount <- starts_joinedLSOA %>%
  st_drop_geometry() %>% 
  group_by(LSOA11CD) %>% 
  summarise(bikecount = n())

LSOAcount2 <- starts_joinedLSOA2 %>%
  st_drop_geometry() %>% 
  group_by(LSOA11CD) %>% 
  summarise(bikecount = n())

LondonLSOA_withBikes <- LondonLSOA %>% 
  left_join(LSOAcount) %>% 
  filter(bikecount < 100000000) %>%
  mutate("Bike rentals per resident (per month)" = bikecount / USUALRES) %>%
  relocate(LSOA11NM, .before = LSOA11CD)

LondonLSOA_withBikes2 <- LondonLSOA %>% 
  left_join(LSOAcount2) %>% 
  filter(bikecount < 100000000) %>%
  mutate("Bike rentals per resident (per month)" = bikecount / USUALRES) %>%
  relocate(LSOA11NM, .before = LSOA11CD)


# Calculating average travel time

LSOAtraveltime <- starts_joinedLSOA %>%
  st_drop_geometry() %>%
  group_by(LSOA11CD) %>%
  summarise(avg_rentaltime = median(convTravelTime))

LSOAtraveltime2 <- starts_joinedLSOA2 %>%
  st_drop_geometry() %>%
  group_by(LSOA11CD) %>%
  summarise(avg_rentaltime = median(convTravelTime))

LondonLSOA_traveltime <- LondonLSOA %>%
  left_join(LSOAtraveltime) %>%
  filter(avg_rentaltime < 100001) %>%
  mutate("Average rental time (minutes)" = avg_rentaltime) %>%
  relocate(LSOA11NM, .before = LSOA11CD)

LondonLSOA_traveltime2 <- LondonLSOA %>%
  left_join(LSOAtraveltime2) %>%
  filter(avg_rentaltime < 100001) %>%
  mutate("Average rental time (minutes)" = avg_rentaltime) %>%
  relocate(LSOA11NM, .before = LSOA11CD)

```

Bike rentals per resident 
===========================


Column {data-width=650}
-----------------------------------------------------------------------

### Total bike rentals June 2019

```{r}
valueBox(nrow(bikes062019), icon = "fa-bicycle")
```

### Total bike rentals June 2020

```{r}
valueBox(nrow(bikes062020), icon = "fa-bicycle")
```

### Increase in bike rentals 

```{r}
valueBox(round(((nrow(bikes062020) / nrow(bikes062019)) - 1) * 100, 2), icon = "fa-percent")
```

Column {data-width=1000}
-----------------------------------------------------------------------

### Bike rentals per resident (June 2019)

```{r}
# Plotting bike rentals per resident 2019 vs 2020

tmap_mode("view")

# 06/2019 Data
#Bbox inner: -0.5,51.27,0.35,51.7
tm_shape(LondonWard, bbox = c(-0.24,51.45,0,51.55)) +
  tm_fill(col = "white", palette = c("lightgrey", "white")) +
  tm_borders() +
  tm_shape(LondonLSOA_withBikes) +
  tm_polygons(col = "Bike rentals per resident (per month)", style = "fixed", breaks = c(0, 0.3, 0.6, 1, 2, 3, 4, 7, 10, 30, 60)) 

```

> Source data from: <https://tfl.gov.uk/info-for/open-data-users/our-open-data?intcmp=3671>

### Bike rentals per resident (June 2020)

```{r}
#06/2020 Data
#Bbox inner: -0.5,51.27,0.35,51.7
tm_shape(LondonWard, bbox = c(-0.24,51.45,0,51.55)) +
  tm_fill(col = "white", palette = c("lightgrey", "white")) +
  tm_borders() +
  tm_shape(LondonLSOA_withBikes2) +
  tm_polygons(col = "Bike rentals per resident (per month)", style = "fixed", breaks = c(0, 0.3, 0.6, 1, 2, 3, 4, 7, 10, 30, 60)) 

```

> Source data from: <https://tfl.gov.uk/info-for/open-data-users/our-open-data?intcmp=3671>

Column {data-width=200}
-----------------------------------------------------------------------

### Usage of bikesharing in London 08/2018 - 09/2020

```{r}
bikerawdata_grouped <- bikerawdata %>%
  mutate(Start_onlyDate = as.Date(Start.Date, "%d/%m/%Y"),isLongRent = (Duration > 1200)) %>%
  group_by(Start_onlyDate, isLongRent) %>% 
  summarise(rentcount = n())

bikerawdata_grouped <- bikerawdata_grouped %>% mutate("Rent is longer than 20 minutes:" = isLongRent)

q <- ggplot(data = bikerawdata_grouped, aes(x = Start_onlyDate ,y = rentcount)) +
  geom_point(size = 1) +
  geom_smooth(method = "gam", se = FALSE) +
  labs(x = "date",
       y = "number of rents") +
  theme(legend.position = "bottom")

ggplotly(q)

```


Average rental time
=============================


Column {data-width=650}
-----------------------------------------------------------------------

### Average rental time June 2019 (minutes)

```{r}
x <- round(sum(LondonLSOA_traveltime$avg_rentaltime) / nrow(LondonLSOA_traveltime),2)
valueBox(x ,icon = "fa-clock")

```

### Average rental time June 2020 (minutes)

```{r}
y <- round(sum(LondonLSOA_traveltime2$avg_rentaltime) / nrow(LondonLSOA_traveltime2),2)
valueBox(as.difftime(y, units = "mins", format = "%M:%S"), icon = "fa-clock")

```

### Increase in average rental time (minutes)

```{r}
z <- round((sum(LondonLSOA_traveltime2$avg_rentaltime) / nrow(LondonLSOA_traveltime2)) - (sum(LondonLSOA_traveltime$avg_rentaltime) / nrow(LondonLSOA_traveltime)), 2)
valueBox(as.difftime(z, units = "mins", format = "%M:%S"), icon = "fa-arrow-up")

```

Column {data-width=1000}
-----------------------------------------------------------------------

### Average rental time (June 2019)

```{r}
# Plotting average rental time 2019 vs 2020
# STYLE ABÄNDERN!

#06/2019 Data
tm_shape(LondonWard, bbox = c(-0.24,51.45,0,51.55)) +
  tm_fill(col = "white", palette = c("lightgrey", "white")) +
  tm_borders() +
  tm_shape(LondonLSOA_traveltime) +
  tm_polygons(col = "Average rental time (minutes)", style = "fixed", breaks = c(5, 7.5, 10, 12.5, 15, 17.5, 20, 25, 30, 40)) 

```

> Source data from: <https://tfl.gov.uk/info-for/open-data-users/our-open-data?intcmp=3671>

### Average rental time (June 2020)

```{r}
#06/2020 Data
tm_shape(LondonWard, bbox = c(-0.24,51.45,0,51.55)) +
  tm_fill(col = "white", palette = c("lightgrey", "white")) +
  tm_borders() +
  tm_shape(LondonLSOA_traveltime2) +
  tm_polygons(col = "Average rental time (minutes)", style = "fixed", breaks = c(5, 7.5, 10, 12.5, 15, 17.5, 20, 25, 30, 40))

```

> Source data from: <https://tfl.gov.uk/info-for/open-data-users/our-open-data?intcmp=3671>

Column {data-width=200}
-----------------------------------------------------------------------

### Usage and rent duration of bikesharing in London 08/2018 - 09/2020

```{r}
p <- ggplot(data = bikerawdata_grouped, aes(x = Start_onlyDate ,y = rentcount)) +
  geom_point(size = 1, alpha = 0.3, aes(color = `Rent is longer than 20 minutes:`)) +
  geom_smooth(method = "gam", se = FALSE, aes(color = `Rent is longer than 20 minutes:`)) +
  labs(x = "date",
       y = "number of rents") +
  theme(legend.position = "bottom")

ggplotly(p)

```



Bike rental stations 
=========================

Column {data-width=650}
-----------------------------------------------------------------------

### Average bike rentals per station (June 2020)

```{r}
valueBox(round(sum(sum_startpoints2$stationcount) / nrow(sum_startpoints2),2), icon = "fa-bicycle")
```

### Highest amount of rentals at one rental station (Hyde Park Corner, June 2020)

```{r}
valueBox(9520, icon = "fas fa-landmark")
```

Column {data-width=1000}
-----------------------------------------------------------------------

### Bikes rentals per station (June 2020)

```{r}
sum_startpoints2 <- sum_startpoints2 %>%
  mutate("Total bike rentals" = stationcount) %>% 
  st_as_sf()

tm_shape(sum_startpoints2, size = 0.15) +
  tm_dots(size = "Total bike rentals", col = "Total bike rentals", style = "fixed", breaks = c(0, 1000, 2000, 3000, 4000, 6000, 8000, 10000))

```

> Source data from: <https://tfl.gov.uk/info-for/open-data-users/our-open-data?intcmp=3671>



Exemplary journey of bike 15776
=========================

Column {data-width=650}
-----------------------------------------------------------------------

### Number of Bike 15776's rentals from 08/2018 to 09/2020

```{r}
valueBox(dim(bike_15776)[1], icon = "fa-bicycle")
```

### Bike 15776's average rental duration

```{r}
valueBox(sprintf("%02dh %02dm", seconds_to_period(mean(bike_15776$Duration))@hour, minute(seconds_to_period(mean(bike_15776$Duration)))), icon = "fa-clock")
```

### `r paste("Bike 15776's longest rental duration on", format(as.POSIXct("29/07/2019 18:09", format = "%d/%m/%Y %H:%M"), "%B %d, %Y"))`

```{r}
valueBox(sprintf("%02dh %02dm", seconds_to_period(max(bike_15776$Duration))@hour, minute(seconds_to_period(max(bike_15776$Duration)))), icon = "fa-clock")
```

Column {data-width=1000}
-----------------------------------------------------------------------

### Map of all rentals of Bike 15776

```{r}
tm_shape(allroutes) +
  tm_lines(alpha = 0.1,
           col = "darkblue",
           palette = c("darkblue", "orange"),
           lwd = 2)
```

> Source data from: <https://tfl.gov.uk/info-for/open-data-users/our-open-data?intcmp=3671>

